### 2020-06-10 explanatory text.Rmd
explanatory text

---
title: "Application Outline"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Application aims

The aim of this application is to provide a statistical foundation for Bateson challenges on the basis of race and gender. Building on the work of Kadane^[Kadane, 2018. _Statistics for Bateson Challenges_, **17**, 1-13.] and Barrett^[Barrett, 2007. _The American Statistician_, **61**, 4, 296–301.], this application makes use of _Bayesian statistical methods_ to identify potential baises in jury selection, while taking into account historical (from previous trials) jury selection behavior by both prosecutor and defense attorneys.

## Components of the analysis

This application relies on two components. The **first component** is historical jury selection data (stored in a compiled database) that is used to inform our 'prior expectations' about jury selection behavior by a given attorney. The **second component** is the jury selection pattern observed in the trial at hand (and entered into this application). The combination of these two elements allow us to make determinations of bias, based on the current trials data that are _conditioned_, or accounting for, historical jury selection practices. Combining information from the current trial with historical information is the essence of the Bayesian analysis: we have updated our prior beliefs based on the new data presented to us.

## Interpreting the results

Looking at the plots above, we can see this playing out. In these plots $d$ is our measure of bias: **positive** values of $d$ reflect bias _against_ the cognizable class, while **negative** values  of $d$ reflect bias _for_ the cognizable class. If $d$ is **zero**, then jury selection is _neutral_.  Given that we are interested in accounting for uncertainty (or, our degree of confidence) in our estimates, we represent $d$ as a distribution rather than a single value.

The greyed-out distribution in the background represents our 'prior expectation' of what the value of $d$ should be based on historical trial data. Here we have one prior expectation for prosecutors and for the defense. The blue (for prosecution) and red (for the defense) distributions in the foreground represent our _updated expectations_ based on the data entered into the application. This updated information is simply the product of the prior information and the information contained in the table on the left. The dashed lines indicate _credible intervals_^[These are 95% credible intervals, i.e., they contain 95% of the values of $d$.] for each of our distributions of $d$. These intervals indicate the regions of the distributions that we have a high degree of confidence contain the true values of $d$. If these credible intervals do not contain zero, this suggests evidence of bias for or against the cognizable class. Alternatively, if these credible intervals do contain zero, we have less confidence that the true value of $d$ is not, in fact zero, and so have no evidence of bias.