### 2020-06-17 prototype/explanatory text.Rmd
revisions, renaming bias parameter as "b"

---
title: "About"
output: 
  github_document:
       pandoc_args: --webtex
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This application estimates how likely an attorney used peremptory challenges during jury selection to strike potential jurors based on their race or gender. The analysis builds on prior work by @Kadane2018 and @Barrett2007 on statistical analysis for _Batson_ challenges.

To estimate such bias, this application relies on (1) user-inputted data on how an attorney has used strikes in the case before them; and (2) if available, jury-selection data on that attorney's use of strikes in some past cases during 2013-2017 in the federal district court for the District of Connecticut.

Based on this data, this application graphs $b$ -- our measure of bias -- for the prosecution and the defense: **positive** values reflect bias _against_ the cognizable class, while **negative** values reflect bias _for_ the cognizable class. If $b = 0$, that there is zero bias.

The plots show how likely each possible value of $b$ is, given the data we have. The curve's spread (the distribution's *variance*) depicts how confident we should be in any particular value of $b$ from the data we have.

The dashed lines indicate a pre-set _credible interval_, i.e., here the values of $d$ between which 80% of the likely values of $b$ fall. If zero falls within this interval, then we should be wary about (have low confidence in) inferring bias (for or against the cognizable class) from the strike data alone. If zero falls outside this interval, we should feel confident in inferring bias from that strike data.

The grey graph depicts $b$ based _only_ on a pre-set distribution of $b$ to reflect a starting point that bias is unlikely, as adjusted by any historical strike data we have for a particular selected attorney. This represents our *prior beliefs* about an attorney's possible bias in using peremptory challenges, before we know anything about how any attorney uses peremptory strikes in the case before us. The red (for prosecution) and blue (for the defense) distributions in the foreground represent our _updated_ expectations after accounting for the user-inputted strike data.

### References