### 2020-06-22 prototype/model_text.Rmd
added a seperate "model" explanation -- currently incomplete, needs likelihood function; simplified "about" text.

---
title: "Model"
output: 
  github_document:
       pandoc_args: --webtex
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Like @Kadane2018 and @Barrett2007, we model the probability that the struck person belonged (or not) to the cognitive class as a function of (1) the number of cognizable class members (e.g., racial minorities, women) and non-class members (e.g., White people, men) in the set of people who could have been struck; and (2) an unobserved variable $b$ that indicates how much weight the striking party placed on the struck juror's membership in the cognitive class (e.g., whether that juror was racial minority or White; women or man).

For any given jury selection $i$, let $j$ denote a peremptory strike used (among $t$ total strikes used), and let $X_{ij}$ denote whether a party used that strike on a person who belongs to a "cognizable class", _i.e._, $X_{ij} = 0$ if no, and $X_{ij} = 1$ if yes. If "race" is the bias type of interest, the cognizable class ($X = 1$) is racial minority jurors ($X = 0$ for White jurors).  If "gender" is the bias of interest, the cognizable class is female jurors ($X=0$ for male jurors).

We are interested in estimating the likely values of $b$, given the strike data we have. There is zero bias only if the probability of striking a cognizable class member is the same as the probability of striking a non-cognitive class member.  Accordingly,

$$w = \frac{Pr(X=1)}{Pr(X = 0)}\\
w = e^b$$

This way, if it is equally likely that a member and non-member of the cognizable class will be stricken, then $w = \frac{Pr(X=1)=0.50}{Pr(X = 0)=0.50}=1$, then our measure of bias is zero, because if $w=e^b$, and $e^b=1$ then $b=log(1)=0$. If $b>1$, we infer that the the party has bias favoring a strike against a juror falling within the cognitive class ($X=1$, e.g., the juror is a racial minority). Where $b<1$, the party has bias toward striking a juror falling outside the cognitive class ($X=0$, e.g., the juror is White).

In turn, let $c$ denote the number of cognizable class members; and $m$ denote the number of cognizable class non-members, such that $c+m$ is the total number of jurors potentially subject to strike.

$$Pr(X_{ij}) =
\begin{cases}
  \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }X=1\\    
  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }X=0  
\end{cases}
$$

### References