### 2021-07-26 analysis/paper_draft.Rmd
added motivation for power prior; minor edits

---
title: "Batson Challenges Using Historical Selection Data"
date: "`r Sys.Date()`"
bibliography: references.bib
author:
  - name: "Sachin S. Pandya"
    affiliation: "University of Connecticut"
    orcid_id: 0000-0001-7387-1307
  - name: "Xiaomeng Li"
    affiliation: "University of Connecticut"
  - name: "Timothy E. Moore"
    affiliation: "University of Connecticut"
    orcid_id: 0000-0002-9576-0517
output: distill::distill_article
---

# Introduction

This paper demonstrates how to use contemporaneous and historical jury-selection data to estimate the probability that an attorney has used peremptory challenges to strike potential jurors because of juror race or gender, in violation of _Batson v. Kentucky_ (1986), and its progeny.

# Background and Related Literature

## Jury Selection

In the US, once a group of individuals selected for jury-service appear in court as scheduled, this group is questioned by the parties' attorneys and the trial judge.  During this "voir dire" process, each prospective juror may be (1) excused by the trial judge for cause; (2) selected for the jury; or (3) dismissed upon the exercise by one of the parties' attorneys of a peremptory challenge. The key feature of a peremptory challenge is that, with it, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury, but without the burden of explaining why. Each side is assigned a number of peremptory challenges and may use them to strike any potential juror without explanation. The number of peremptory challenges varies by jurisdiction, as does peremptory-challenge procedure, including the order of strikes and the number of peremptory strikes a party may exercise at a time [@nscs-jury2018; @Williams2017].

<!--
Courts also vary as to the number of peremptory strikes a party exercises at a time [@Williams2017, p. 297-298]. This particularly matters where each side does not have the same number of strikes.  For example, if the prosecution has six strikes and the defendant has ten, the custom may be to allow one prosecutor strike, then _two_ defense strikes, then the second prosecutor strike, then two more defense strikes, and so on, until all the strikes are exercised.
-->

Under _Batson v. Kentucky_ (1986)_, however, a party's exercise of a peremptory challenge violates the Equal Protection Clause of the US Constitution if motivated by a prospective juror's race, gender, or certain other protected characteristics. When raising a _Batson_ challenge during jury selection, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status. Once established, the challenged party bears the burden of producing one or more nondiscriminatory reasons for that strike.  Thereafter, the challenger bears the burden of persuading the trial judge that the race or gender, not the proffered nondiscriminatory reason, was more likely than not the real motivation for the challenged strike(s). For discussion, see [@LaFave2020, sec. 22.3]. Some States have adopted variants of _Batson_ that reduce or eliminates the challenger's burden to establish a prima facie case^[E.g., State v. Holloway, 209 Conn. 636, 646 & n.4 (1989); State v. Chapman, 317 S.C. 302, 305-06 (1995); see also State v. Antwine, 743 S.W.2d 51, 64 (Mo. 1987) (trial judges must consider prosecutor's explanation to determine whether defendant established a prima facie case).]; and, more recently, replaced a focus on the motive for strikes with a query into whether a hypothetical "objective observer" would find that race or ethnicity was a "factor" in the use of strikes.^[Wash. R. 37(e)].

By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on peremptory challenges in past cases. For example, in a _Batson_ challenge against a prosecutor in a criminal case, such historical data would be relevant to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in that case.^[See, e.g., Miller-El v. Dretke, 545 U.S. 231, 253 (2005); Riley v. Taylor, 277 F.3d 261, 280-81, 283 (3d Cir. 2001); Cochran v. Herring, 43 F.3d 1404, 1412 (11th Cir. 1995); see also Flowers v. Mississippi, 139 S. Ct. 2228, 2243 (2019) (kinds of evidence to support Batson challenge includes "relevant history of the State's peremptory strikes in past cases")]

Despite _Batson_ and its progeny, "prosecutors in criminal cases, and defense attorneys in civil case are more likely to use their challenges to dismiss African American jurors, whereas criminal defense attorneys and civil plaintiff's attorneys are more likely to exercise their strikes on white prospective jurors." [@Diamond2018]. This pattern persists, despite the fact that both sides' attorneys can often observe the race and gender of those in the venire who were and were not struck by peremptory challenge. The research literature suggests several explanations. These include a reluctance to raise a _Batson_ challenge on the premise that such a challenge raises with it the embarrassment, discomfort, and reputation costs with being challenged as racist [@OBrien2019, 29; @Offit2021]. Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge. Moreover, drawing inferences from historical data -- the use of peremptory challenges in _past_ cases -- is hard, because jury-selection records are not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise require painstaking investigation [e.g., @Craft2018].

In theory, however, were an attorney's strike history in past cases easily available, such strike history can be compared to what we would expect to happen if, following the same procedure, venire members were struck at random [@Kadane2018; @Kadane2018a]. To do so, one would need to, for any peremptory-selection method, a corresponding likelihood function for how a prosecutor or defense attorney strikes prospective jurors of a protected class (e.g., racial-minorities). Yet, even with agreement on the appropriate statistical analysis, it is hard to complete the analysis in time to matter, unless "statisticians accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]

# Discussion

Here, we discuss how to assess whether one can credibly infer bias, for Batson purposes, based on an attorney's pattern of peremptory challenges in a particular case either alone or in combination with historical data on that attorney's pattern of strikes in past cases. In doing so, we make two distinctive contributions.

First, we use the power prior to control how much the strike history in past cases affects the posterior distribution of the bias parameter $b$. The power prior is a way to create informative prior distributions based on historical data so as to control how much historical data affects the posterior distribution of the unobserved variables of interest. The key insight is to raise the likelihood function based on the historical data to a power parameter. For a review, see @Ibrahim2015.

Second, we implement the analysis as prototype software that an attorney could use in real time during jury selection to assess how likely an attorney striking prospective jurors in violation of _Batson_.

# Model

Under the "struck jury" method of peremptory challenges, the trial judge rules on all challenges for cause before the parties exercise any peremptories, so that the number of potential jurors who remain is the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides. Accordingly, in a non-capital felony trial in federal district court using this method, the trial judge will, after excusing jurors for cause, make sure that there are 28 potential jurors remaining before the parties exercise their peremptory challenges (12 jurors plus 16 peremptories, assuming no alternate jurors). Then, the parties exercise their strikes on those 28 potential jurors.

In a typical variant of the "struck jury" method, the parties strike in an alternating sequence, with the prosecution striking first. Where the number of strikes are not the same for the parties, an alternating sequence may result in one party exercising multiple strikes per turn. If a party decides not to use all its strikes, the jurors are seated based on the order in which they were selected or at random. Under "strike-and-replace" or "jury box" method, enough potential jurors are seated to fill the jury box. Some of them are excused for cause, and the parties decide whether or not to strike any of the remaining jurors with their peremptory challenges.  Once that is completed, new potential jurors take the seats of the excused or stricken jurors, and the process repeats until all the jury seats are filled.  At any point, strikes can be used on anyone in the jury box, including those there previously but not struck.

Following @Kadane2018 and @Barrett2007, we model these peremptory-challenge processes as follows. For any given jury selection $i$, let $j$ denote a peremptory strike used, and let $\delta_{ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class". If "race" is the bias type of interest, the cognizable class is racial minority jurors ($\delta_{ij} =1$, 0 for White jurors).  If "gender" is the bias of interest, the cognizable class is female jurors ($\delta_{ij} = 1$, 0 for male jurors).

In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike.

If there is no bias, the probability is $\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking a cognizable class member, and $1-\frac{c_{ij}}{c_{ij}+m_{ij}}$. If we define the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$ by adding one parameter $w$, we can measure the bias by different values of $w$. When $w=1$, it corresponds to the situation when there is no bias. When $w>1$, the probability of striking a cognizable class member is greater than non cognizable class, which means a discriminatory against cognizable class. On the contrary, when $w<1$, the probability of striking a cognizable class member is smaller than non cognizable class, which means a preference of cognizable class.

However, under this set up, the weight of the non-cognizable class is the reciprocal of the weight of cognizable class, which is not a desirable character. Using $b = log(w)$ could avoid this issue. For any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation \@ref(eq:model1)

\begin{equation}
Pr(\delta_{ij} | b) = 
\begin{cases}
  \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=1\\    
  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  
\end{cases} (\#eq:model1)
\end{equation}

This is equivalent to

$$Pr(\delta_{ij}| b) = (\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}$$

Given the strike data we have, i.e., $\delta_{ij}$, $c_{ij}$, and $m_{ij}$, by estimating the value of $b$, we can measure whether there is bias when striking potential jurors.

If $b = 0$, there is no bias, and the probability of strike is simply a function of the share of cognizable members (non-members) in the pool of prospective jurors that could be struck. If $b>0$, we infer that the the party has bias favoring a strike against a juror falling within the cognizable class (e.g., the juror is a racial minority). Where $b<0$, the party has a preference for a juror within the cognizable class.

The likelihood function of b is shown in equation \@ref(eq:model2)

\begin{equation}  
L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}} (\#eq:model2)
\end{equation}
<!-- \[b \sim \text{Normal}(0,2)\] -->

where $n_{i}$ is the total number of jury selections; and $n_j$ is the total number of peremptory strikes. In @Kadane2018, $b$ is estimated under Bayesian framework with a uniform prior from -6 to 6. In this paper, we also would like to incorporate the historical information. In addition, we want to allow different weight for historical information, considering people's preference might change over time. For instance, we can set a smaller weight if the historical data is collected many years ago than recent information. Thus we use power prior to incorporate the historical information [@Ibrahim2000; @Ibrahim2015]. The formulation of the power prior is shown in equation \@ref(eq:model3)

\begin{equation}
\pi(b|D_0,a_0) \propto L(b|D_0)^{a_0}\pi_0(b) (\#eq:model3)
\end{equation}

where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information, and $D_0$ is the observed historical data. $L(b|D_0)$ is the likelihood function of $b$ given the historical data. $\pi_0(b)$ is the initial prior before the historical data is observed. In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.

<!-- The prior distribution for $b$ ($\sim \text{Normal}(0,2)$] is justified on the grounds that the party bringing the Batson challenge bears the ultimate burden of persuading the trial judge of an illegal use of strikes. -->

The posterior distribution of $b$ after including the historical information through the power prior is proportional to the product of likelihood function of $b$ and the power prior of $b$ as shown in \@ref(eq:model4)

\begin{equation}
L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8}) (\#eq:model4)
\end{equation}

where $L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{\delta_{0ij}} (\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{1-\delta_{0ij}}$ is the likelihood function of $b$ given historical data. $n_{0i}$ is the total number of jury selections in the historical data, and $n_{0j}$ is the total number of peremptory strikes in the historical data. $\delta_{0ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class" in the historical trials.

The posterior distribution does not have a closed form.  Accordingly, we used the Metropolis-Hasting algorithm sample $b$ from the posterior distribution. Traceplots, Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998} and Geweke’s diagnostics [@Gewe1992] were used to measure the convergence. 
<!--

# Model 2 - Sequential Strike Method

Under the "sequential strike" method of peremptory challenges, the trial judge rules on all challenges for cause before the parties exercise any peremptory challenges. Then, the remaining prospective jurors appear one at a time before the attorneys for prosecutor and defendant, and those attorneys alternate as the order in which they decide whether to strike or not the prospective juror before them. For example, if the prosecutor had first choice whether to strike juror #1, the defendant's attorney would have first choice whether to strike juror # 2. If neither party exercises a strike on the prospective juror before them, that prospective juror is assigned a seat on the jury and the parties' attorneys cannot thereafter strike that juror. The process continues until the parties run out of strikes or the number seats on the jury (including alternates) have been filled.

Accordingly, we model the sequential-strike process as follows.

[insert statistical model]
-->

# Simulation

We conducted a simulation study to measure model performance under different values of the bias parameter, different amount of historical information and whether the bias/preference changes over time. To do this, we generated data using seven different values of the bias parameter $b = {-3,-2,-1,0,1,2,3}$. As introduced above, $b<0$ represents a preference for a juror within the cognizable class, and $b>0$ represents a bias against the cognizable class. A greater absolute value of $b$ represents greater preference or bias for the cognizable class, while $b=0$ represents no bias. We also considered three sizes of historical data, the same size as current data, double size of current data and triple size of current data. We also considered seven different values of the bias parameter for generating the historical data and allow the bias parameter of historical data to be different from current data. The different value of bias parameter from historical data to current data represents a change of preference for cognizable class.

By using the power prior to incorporate the historical information, we can control the weight that we put on historical data by modifying the weight parameter $a$.  We may wish to exercise such control and $a < 1$ if some of the historical data is missing not at random. For example, for any given attorney, there may be only incomplete information available as to that attorney's exercise of peremptory challenges in previous trials. Possible reasons include limited access to the paper records containing such information, either due to rules governing who may access such records or the expense incurred in obtaining copies of such records. Standard jury-selection forms may not always require recording relevant information about prospective jurors (e.g., their race or gender), which may force researchers to combine such records with other data sources (e.g., voter registration rolls) by matching by name.  And even if standard forms or other recordkeeping rules require collecting all relevant information for each past trial, judges, attorneys, and court staff can vary in how much or how often they actually fill in all the required information.

Accordingly, in the simulation study, we tried different values of the weight parameter $a = (0.1,0.3,0.5,0.7,1)$ to show how that weight parameter affected modeling accuracy. We consider 735 scenarios in total, i.e., the combination of 7 different bias parameter of current data and of historical data, 3 different amount of historical data and 5 different values of the power prior weight parameter $a$. For each scenario, we generate 1000 datasets and fit the model on the 1000 generated dataset. We record the mean value of the posterior mean and credible intervals of one thousand model fits for each scenario.

For the credible interval, we set three significance levels (80\%, 90\% and 95\%) and recorded the corresponding coverage rate. For instance, the 90\% coverage rate means the proportion of the 1000 model fits that the credible interval shows the true preference as we set. That is, if we set $b>0$, we calculate the proportion of the 90\% credible intervals that lie on the right of 0 among the 1000 model fits.

The following figures depict the coverage rate under different scenarios. Figure \@ref(fig:b80) depicts the results of significance of estimated $b$ under 80\% credible interval. Figure \@ref(fig:b90) shows the results under 90\% credible interval and figure \@ref(fig:b95) shows the results under 95\% credible interval.

```{r b80,  echo=FALSE, fig.cap="80% significance of bias parameter", out.width = '100%'}

library(ggplot2)

# INSERT simulation and plot code

load_path <- here::here("sim_resultc.csv")

sims <- readr::read_csv(load_path)
sims$a.f <- as.factor(sims$a)
sims$sig <- ifelse(sims$L80 <0 & sims$U80 >0 , "ns", 1) 
sims$sig <- ifelse(sims$L80 >0 & sims$U80 >0 , 
                                    "pos", sims$sig) 
sims$sig <- ifelse(sims$L80 <0 & sims$U80 <0 , 
                                    "neg", sims$sig) 
sims$sig.n <- 0
sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)

p = ggplot()+
  geom_tile(data = sims, aes(x = d, y = d_h,
                        fill = as.factor(sig.n)),
            color="black")+
  facet_grid(h~a)+
  # scale_fill_gradient2(expression(hat(d)), 
  #                      low = "darkred", mid = "white", 
  #                      high = "blue", 
  #                      breaks = c(-3, -2, -1, 0, 1, 2, 3), 
  #                      limits = c(-3.2, 3.2),
  #         labels = c("Neg bias", "", "", "NS","", "", "Bias"))+
  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
  xlab("Current b")+
  ylab("Historical b")+
  scale_fill_manual(values = c("white", "blue"))+
  coord_equal()+
  theme_classic()+
  theme(axis.text = element_text(size =10), 
        axis.title = element_text(size = 16), 
        legend.title.align = 0.05, 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14))
p

```

```{r b90,  echo=FALSE, fig.cap="90% significance of bias parameter", out.width = '100%'}

# INSERT simulation and plot code
sims$a.f <- as.factor(sims$a)
sims$sig <- ifelse(sims$L90 <0 & sims$U90 >0 , "ns", 1) 
sims$sig <- ifelse(sims$L90 >0 & sims$U90 >0 , 
                                    "pos", sims$sig) 
sims$sig <- ifelse(sims$L90 <0 & sims$U90 <0 , 
                                    "neg", sims$sig) 
sims$sig.n <- 0
sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)

p = ggplot()+
  geom_tile(data = sims, aes(x = d, y = d_h,
                        fill = as.factor(sig.n)),
            color="black")+
  facet_grid(h~a)+
  # scale_fill_gradient2(expression(hat(d)), 
  #                      low = "darkred", mid = "white", 
  #                      high = "blue", 
  #                      breaks = c(-3, -2, -1, 0, 1, 2, 3), 
  #                      limits = c(-3.2, 3.2),
  #         labels = c("Neg bias", "", "", "NS","", "", "Bias"))+
  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
  xlab("Current b")+
  ylab("Historical b")+
  scale_fill_manual(values = c("white", "blue"))+
  coord_equal()+
  theme_classic()+
  theme(axis.text = element_text(size =10), 
        axis.title = element_text(size = 16), 
        legend.title.align = 0.05, 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14))
p
```

```{r b95,  echo=FALSE, fig.cap="95% significance of bias parameter", out.width = '100%'}

# INSERT simulation and plot code
sims$a.f <- as.factor(sims$a)
sims$sig <- ifelse(sims$L95 <0 & sims$U95 >0 , "ns", 1) 
sims$sig <- ifelse(sims$L95 >0 & sims$U95 >0 , 
                                    "pos", sims$sig) 
sims$sig <- ifelse(sims$L95 <0 & sims$U95 <0 , 
                                    "neg", sims$sig) 
sims$sig.n <- 0
sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)

p = ggplot()+
  geom_tile(data = sims, aes(x = d, y = d_h,
                        fill = as.factor(sig.n)),
            color="black")+
  facet_grid(h~a)+
  # scale_fill_gradient2(expression(hat(d)), 
  #                      low = "darkred", mid = "white", 
  #                      high = "blue", 
  #                      breaks = c(-3, -2, -1, 0, 1, 2, 3), 
  #                      limits = c(-3.2, 3.2),
  #         labels = c("Neg bias", "", "", "NS","", "", "Bias"))+
  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
  xlab("Current b")+
  ylab("Historical b")+
  scale_fill_manual(values = c("white", "blue"))+
  coord_equal()+
  theme_classic()+
  theme(axis.text = element_text(size =10), 
        axis.title = element_text(size = 16), 
        legend.title.align = 0.05, 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14))
p
```

_Historical b_ means the bias parameter for the generated historical data, current $b$ means the bias parameter for the current data. The values on the right side represent the size of the historical data. The values on the top of the plots represent the weight parameter of the power prior. Blue means the credible interval of $b$ does not include zero, that is $b$ is significant.

From the plot, our model can capture strong bias and bias, when the bias parameter being -3,-2,2,3. When bias is moderate ($b = \{1, -1\}$), the model can detect it when we have historical data with the same direction bias parameter. If the bias parameter of historical data is in an opposite direction, we might get a wrong estimate of current bias parameter.

# The App Prototype

The posterior probability distribution is estimated with `r R.version$version.string'[@R]. The app was built using the shiny package [@shiny].

# Discussion

# Conclusion

# CRediT Author Statement

Sachin S. Pandya: Conceptualization; Xiaomeng Li: Conceptualization; [insert]; Timothy Moore: Conceptualization; [insert]. Based on categories in https://casrai.org/credit/].

# References