### 2021-08-02 analysis/paper_draft.Rmd
updated section on prior literature; edits to main text

---
title: "Batson Challenges Using Historical Data"
date: "`r Sys.Date()`"
bibliography: references.bib
author:
  - name: "Sachin S. Pandya"
    affiliation: "University of Connecticut"
    orcid_id: 0000-0001-7387-1307
  - name: "Xiaomeng Li"
    affiliation: "University of Connecticut"
  - name: "Timothy E. Moore"
    affiliation: "University of Connecticut"
    orcid_id: 0000-0002-9576-0517
output: distill::distill_article
---

# Introduction

This paper shows how to use historical jury-selection data to estimate how likely an attorney has used peremptory challenges to strike potential jurors because of juror race or gender in violation of _Batson v. Kentucky_ (1986), and its progeny.

In the US, once a group of individuals selected for jury-service appear in court as scheduled, this group is questioned by the parties' attorneys and the trial judge.  During this "voir dire" process, a prospective juror, if not selected for the jury or excused by the trial judge for cause, may be dismissed if a party's attorney uses a peremptory challenge against them.

# Background and Related Literature

## Jury Selection

With a peremptory challenge, a party declares a prospective juror ineligible ("strikes" that juror) for a seat on the jury without the burden of explaining why. While countries such as the United Kingdom and Canada have abolished peremptory strikes (Criminal Justice Act 1988, s 118(1); S.C. 2019, c. 25, s 269), federal and State law in the United States continue to authorize their use during jury selection in criminal and civil cases. In the US, peremptory-challenge procedure can vary not only by State, but also by court and judge, including the number of peremptory challenges assigned to each side, the order of strikes, and the number of peremptory strikes a party may exercise at a time [@nscs-jury2018; @Williams2017].

Under the "struck jury" method of peremptory challenges, the trial judge rules on all challenges for cause before the parties exercise any peremptories, so that the number of potential jurors who remain is the sum of number of seats on the jury (plus alternates, if any) and the number of peremptory challenges allotted to both sides.

Then, the parties exercise their strikes on those potential jurors. In one procedure, the parties submit their strikes simultaneously. In another variant, the parties strike in an alternating sequence, with the prosecution striking first. Where the number of strikes are not the same for the parties, an alternating sequence may result in one party exercising multiple strikes per turn. If a party decides not to use all its strikes, the jurors are seated based on the order in which they were selected or at random.

Under a "strike-and-replace" or "jury box" method, enough potential jurors are seated to fill the jury box. Some of them are excused for cause, and the parties decide whether or not to strike any of the remaining jurors with their peremptory challenges.  Once that is completed, new potential jurors take the seats of the excused or stricken jurors, and the process repeats until all the jury seats are filled.  At any point, strikes can be used on anyone in the jury box.

<!--

Note: Constitutionality of Canada's abolition of strikes upheld in https://decisions.scc-csc.ca/scc-csc/scc-csc/en/item/18932/index.do

Courts also vary as to the number of peremptory strikes a party exercises at a time [@Williams2017, p. 297-298]. This particularly matters where each side does not have the same number of strikes.  For example, if the prosecution has six strikes and the defendant has ten, the custom may be to allow one prosecutor strike, then _two_ defense strikes, then the second prosecutor strike, then two more defense strikes, and so on, until all the strikes are exercised.

-->

This paper focuses on _Batson_ challenges to a party's use of peremptory challenge. Under the US Supreme Court's decision in _Batson v. Kentucky_ (1986)_, a party's exercise of a peremptory challenge against one or more prosepctive jurors violates the Equal Protection Clause of the US Constitution if motivated by such jurors' race, gender, or certain other protected characteristics.

For a _Batson_ challenge, the challenging party must typically establish a "prima facie case", that is, enough evidence to infer that a party struck a juror because of such juror's race, gender, or other protected status. If so, the challenged party must produce one or more nondiscriminatory reasons for that strike. Thereafter, the challenger must persuade the trial judge that prospective jurors' race or gender, not the proffered nondiscriminatory reason(s), was more likely than not the real motivation for the challenged strike(s) [@LaFave2020, sec. 22.3]. For similar challenges under State law, some States have reduced or eliminated the challenger's burden to establish a prima facie case [e.g., State v. Holloway, 209 Conn. 636, 646 & n.4 (1989); State v. Chapman, 317 S.C. 302, 305-06 (1995)], while other States have shifted the inquiry away from the striking party's motives to whether a hypothetical "objective observer" would find that race or ethnicity was a "factor" in that party's use of strikes [e.g., Wash. R. 37(e)].

Despite _Batson_ and its progeny, "prosecutors in criminal cases, and defense attorneys in civil case are more likely to use their challenges to dismiss African American jurors, whereas criminal defense attorneys and civil plaintiff's attorneys are more likely to exercise their strikes on white prospective jurors." [@Diamond2018]. Possible explanations include attorney fear that raising a _Batson_ challenge raises with it the embarrassment, discomfort, and reputation costs with being challenged as racist [@OBrien2019, 29; @Offit2021]. Similarly, a trial judge deciding a _Batson_ challenge may similarly be reluctant to find against opposing counsel, particularly if that prosecutor repeatedly appears before that judge.

By law, evidence supporting a _Batson_ challenge may include not only the circumstances of the jury selection itself, but also data on the pattern of peremptory challenges in past cases. For example, in a _Batson_ challenge against a prosecutor in a criminal case, such historical data would be relevant to whether (1) to conclude that the defendant has a prima facie case; (2) to believe a prosecutor from that office who has offered a race-neutral explanation for striking, say, an African-American juror in that case; and (3) to ultimately accept or reject a Batson challenge in that case [e.g., Miller-El v. Dretke, 545 U.S. 231, 253 (2005); Riley v. Taylor, 277 F.3d 261, 280-81, 283 (3d Cir. 2001); Cochran v. Herring, 43 F.3d 1404, 1412 (11th Cir. 1995)]. The US Supreme Court recently identified the kinds of evidence to support _Batson_ challenge as including "relevant history of the State's peremptory strikes in past cases" [e.g, Flowers v. Mississippi, 139 S. Ct. 2228, 2243 (2019)].

Prior work on statistical methods for _Batson_ challenges, however, have largely focused on strikes used in a single case taken alone. For example, @Barrett2007 modeled bias detection for _Batson_ challenges, given a "struck-jury" procedure in which the parties took turns striking prospective jurors. @Gastwirth2016 applied Fisher's exact test to detect bias in the pattern of strikes in each of ten published cases. For each of three cases, @Kadane2018 compared various frequentist approaches (including Fisher's exact test) with a Bayesian approach.

When considering strikes in multiple trials in the same State or judicial district, @Gastwirth2013 criticized the use of a paired t-test to compare the differences in strike rates of Black and non-Black venire members in the aggregate. Rather, both they and @Gastwirth2014 argued for applying the Cochran–Mantel–Haenszel test by taking the difference of the sum of the observed numbers of Black venire members challenged in each trial and the corresponding sum of the expected numbers of Blacks challenged, and then comparing it to the standard deviation of the difference.

Moreover, drawing inferences from peremptory challenges in past cases is hard, because jury-selection records are often not readily available and largely on paper, not electronic or machine-readable formats [@Grosso2017; @Wright2018], and otherwise require painstaking efforts to collect [e.g., @Craft2018].

This may matter for how much weight to assign such historical data if data on the use of strikes in past trials is missing not at random. For example, for any given attorney, there may be only incomplete information on that attorney's use of strikes in previous trials in a particular State or court. Possible reasons include limited access the relevant records (e.g., jury-selection records kept by a jury administrator, not in a case file) and the expense of obtaining copies of them.  Moreover, because some standard jury-selection forms do not require recording relevant juror information, such as juror race or gender, lawyers and researchers must turn to other data sources (e.g., voter registration rolls) with that information. Finally, regardless of what is supposed to be recorded, judges and court staff vary in how often they actually fill in standard forms with the required information.

Finally, even with agreement on the appropriate statistical analysis, it is often hard to complete the analysis in time to matter, unless "statisticians accompany attorneys in the jury selection process or a convenient programme or app is developed." [@Kadane2018, 12]

# Discussion

Here, we extend the Bayesian approach to _Batson_ challenges to assess whether one can credibly infer illegal bias based solely on an attorney's pattern of peremptory challenges using both present strike data and historical data that attorney's strikes in past cases.

In short, our workflow is this: First, specify a model of the peremptory-strike process that best approximates the peremptory-strike process in the jurisdiction of interest. Second, assign a prior distribution to the bias parameter $b$ that reflects the legal requirement that the _Batson_ challenger bear the burden of proof. Here, we assume a normal prior for $b$ with mean 0 and standard deviation 2. Third, using the strike data in past trials generated by the same peremptory-strike process, estimate a posterior distribution for $b$.  Finally, using data on the target attorney's strikes in the present case, use that posterior distribution as the prior for $b$ and update.

In so doing, we make two main contributions. First, we use the power prior to control how much the strike history in past cases affects the posterior distribution of the bias parameter $b$. The power prior is a way to create informative prior distributions based on historical data so as to control how much historical data affects the posterior distribution of the unobserved variables of interest. The key insight is to raise the likelihood function based on the historical data to a power parameter [@Ibrahim2000; @Ibrahim2015].

Second, we implement the analysis as prototype software that an attorney could use in real time during jury selection to assess how likely an attorney striking prospective jurors in violation of _Batson_.

# Model

Following @Kadane2018 and @Barrett2007, we model a struck-jury peremptory-challenge process as follows.

For any given jury selection $i$, let $j$ denote a peremptory strike used, and let $\delta_{ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class". If "race" is the bias type of interest, the cognizable class is racial minority jurors ($\delta_{ij} =1$, 0 for White jurors).  If "gender" is the bias of interest, the cognizable class is female jurors ($\delta_{ij} = 1$, 0 for male jurors).

In turn, let $c_{ij}$ denote the number of cognizable class members subject to strike; and $m_{ij}$ denote the number of cognizable class non-members subject to strike, such that $c_{ij}+m_{ij}$ is the total number of jurors potentially subject to strike. If there is no bias, the probability is $\frac{c_{ij}}{c_{ij}+m_{ij}}$ for striking a cognizable class member, and $1-\frac{c_{ij}}{c_{ij}+m_{ij}}$.

By adding one parameter $w$, we can measure the bias by different values of $w$ by defining the probability of a cognizable class member being struck to be $\frac{wc_{ij}}{wc_{ij}+m_{ij}}$. As a result, however, the weight of the non-cognizable class is the reciprocal of the weight of cognizable class. Using $b = log(w)$ avoids this issue. Accordingly, for any given value of the bias parameter $b$, the probability of strike of a member from either class, or $Pr(\delta_{ij})$, is shown in equation \@ref(eq:model1)

\begin{equation}
Pr(\delta_{ij} | b) = 
\begin{cases}
  \frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=1\\    
  \frac{m_{ij}}{(e^b)c_{ij}+m_{ij}} & \text{for }\delta_{ij}=0  
\end{cases} (\#eq:model1)
\end{equation}

This is equivalent to

$$Pr(\delta_{ij}| b) = (\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}}$$

Given the strike data we have, i.e., $\delta_{ij}$, $c_{ij}$, and $m_{ij}$, by estimating the value of $b$, we can measure bias when a party is striking potential jurors. If $b = 0$, there is no bias, and the probability of strike is simply a function of the share of cognizable members (non-members) in the pool of prospective jurors that could be struck. If $b>0$, we infer that the the party has bias favoring a strike against a juror falling within the cognizable class (e.g., the juror is a racial minority). Where $b<0$, the party has a preference for a juror within the cognizable class.

The likelihood function of $b$ is shown in equation \@ref(eq:model2)

\begin{equation}  
L(b|\delta) = \prod_{i=1}^{n_i} \prod_{j=1}^{n_j}(\frac{(e^b)c_{ij}}{(e^b)c_{ij}+m_{ij}})^{\delta_{ij}} (\frac{m_{ij}}{(e^b)c_{ij}+m_{ij}})^{1-\delta_{ij}} (\#eq:model2)
\end{equation}
<!-- \[b \sim \text{Normal}(0,2)\] -->

where $n_{i}$ is the total number of jury selections (trials); and $n_j$ is the total number of peremptory strikes. In @Kadane2018, $b$ is estimated under Bayesian framework with a uniform prior from -6 to 6.

In this paper, we also would like to incorporate the historical data on strikes in past cases and be able to adjust the weight of that historical information on the posterior distribution of the bias parameter.  Accordingly, we introduce the power prior \@ref(eq:model3)

\begin{equation}
\pi(b|D_0,a_0) \propto L(b|D_0)^{a_0}\pi_0(b) (\#eq:model3)
\end{equation}

where $0 \le a_0 \le 1$ is the parameter controlling the weight of the historical information; $D_0$ is the observed historical data; $L(b|D_0)$ is the likelihood function of $b$ given the historical data; and $\pi_0(b)$ is the initial prior before the historical data is observed. In this paper, we assume a normal prior for $\pi_0(b)$ with mean 0 and standard deviation 2.

<!-- The prior distribution for $b$ ($\sim \text{Normal}(0,2)$] is justified on the grounds that the party bringing the Batson challenge bears the ultimate burden of persuading the trial judge of an illegal use of strikes. -->

After including the historical information through the power prior, the posterior distribution of $b$  is proportional to the product of likelihood function of $b$ and the power prior of $b$ as shown in \@ref(eq:model4)

\begin{equation}
L(b|\delta) = L(b|\delta)(L(b|\delta_0))^{a_0}\exp(-\frac{b^2}{8}) (\#eq:model4)
\end{equation}

where $L(b|\delta_0) = \prod_{i=1}^{n_{0i}} \prod_{j=1}^{n_{0j}}(\frac{(e^b)c_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{\delta_{0ij}} (\frac{m_{0ij}}{(e^b)c_{0ij}+m_{0ij}})^{1-\delta_{0ij}}$ is the likelihood function of $b$ given historical data. $n_{0i}$ is the total number of jury selections in the historical data, and $n_{0j}$ is the total number of peremptory strikes in the historical data. $\delta_{0ij}$ denote whether or not a party used that strike on a person who belongs to a "cognizable class" in the historical trials.

The posterior distribution does not have a closed form.  Accordingly, we used the Metropolis-Hasting algorithm to sample $b$ from the posterior distribution. Traceplots, Gelman-Rubin convergence diagnostic [@Gelman1992; @Gelman1998} and Geweke’s diagnostics [@Gewe1992] were used to measure the convergence.

<!--

# Model 2 - Sequential Strike Method

Under the "sequential strike" method of peremptory challenges, the trial judge rules on all challenges for cause before the parties exercise any peremptory challenges. Then, the remaining prospective jurors appear one at a time before the attorneys for prosecutor and defendant, and those attorneys alternate as the order in which they decide whether to strike or not the prospective juror before them. For example, if the prosecutor had first choice whether to strike juror #1, the defendant's attorney would have first choice whether to strike juror # 2. If neither party exercises a strike on the prospective juror before them, that prospective juror is assigned a seat on the jury and the parties' attorneys cannot thereafter strike that juror. The process continues until the parties run out of strikes or the number seats on the jury (including alternates) have been filled.

Accordingly, we model the sequential-strike process as follows.

[insert statistical model]
-->

# Simulation

We conducted a simulation study to measure model performance under different values of the bias parameter, different amount of historical information and whether the bias/preference changes over time.

To do this, we generated strike data for current trial using seven different values of the bias parameter ("current" $b = {-3,-2,-1,0,1,2,3}$. As introduced above, $b<0$ represents a preference for a juror within the cognizable class; and $b>0$ represents a bias against the cognizable class. A greater absolute value of $b$ represents greater preference or bias for the cognizable class, while $b=0$ represents no bias.

We also generated historical strike data of three sizes (same, double, and triple the size of current data) as well as seven different values of the bias parameter for generating the historical data ("historical" $b = {-3,-2,-1,0,1,2,3}$).

We let historical $b$ differ from current $b$ (our parameter of interest) for two reasons. First, it allows us to simulate noise associated with historical strike data that is incomplete not at random. Second, it could also simulate a real difference in the bias parameter that depends on a feature that the current trial has but that _all_ the past trials in the historical data do not. To illustrate, assume a prosecutor with two true bias values: $b > 0$ if the criminal defendant is Black, $b = 0$ if not. If our historical data contains only strike data in cases with non-Black defendants, then that prosecutor's historical $b$ will differ from the current case with a Black defendant.

By using the power prior, we can control the weight that we put on historical data by modifying the weight parameter $a$. If $a = 1$, the historical strike data is equally weighted with the present data on strikes. If $a < 1$, the historical data matters proportionally less. As discussed above, if some of the historical data is missing not at random, we may wish to assign $a < 1$ to evaluate how sensitive the posterior of the bias parameter to the historical data on strikes. Accordingly, in the simulation study, we tried different values of the weight parameter $a = (0.1,0.3,0.5,0.7,1)$ to show how that parameter affected modeling accuracy.

We considered 735 scenarios in total, i.e., the combination of 7 different bias parameters of current data and of historical data, 3 different amounts of historical data and 5 different values of the power prior weight parameter $a$. For each scenario, we generated 1000 datasets and fit the model on those 1000 dataset. We recorded the mean value of the posterior mean and credible intervals of one thousand model fits for each scenario.

In turn, we set three credible interval level (80\%, 90\% and 95\%) and recorded the coverage rates, i.e., the proportion of times those intervals actually contained the true value of the (current) bias parameter $b$. For instance, a 90\% coverage rate means that in 900 of 1000 model fits, a given credible interval contained what we fixed as the value of current $b$. Put another way, if we set $b>0$, we calculate the proportion of the 90\% credible intervals that lie on the right of 0 among the 1000 model fits.

The following figures depict the coverage rates of $b$ under the 80\% credible interval (Figure \@ref(fig:b80)), under the 90\% credible interval (Figure \@ref(fig:b90)), and under the 95\% credible interval (Figure \@ref(fig:b95)).

```{r b80,  echo=FALSE, fig.cap="80% credible interval", out.width = '100%'}

library(ggplot2)

# INSERT simulation and plot code

load_path <- here::here("sim_resultc.csv")

sims <- readr::read_csv(load_path)
sims$a.f <- as.factor(sims$a)
sims$sig <- ifelse(sims$L80 <0 & sims$U80 >0 , "ns", 1) 
sims$sig <- ifelse(sims$L80 >0 & sims$U80 >0 , 
                                    "pos", sims$sig) 
sims$sig <- ifelse(sims$L80 <0 & sims$U80 <0 , 
                                    "neg", sims$sig) 
sims$sig.n <- 0
sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)

p = ggplot()+
  geom_tile(data = sims, aes(x = d, y = d_h,
                        fill = as.factor(sig.n)),
            color="black")+
  facet_grid(h~a)+
  # scale_fill_gradient2(expression(hat(d)), 
  #                      low = "darkred", mid = "white", 
  #                      high = "blue", 
  #                      breaks = c(-3, -2, -1, 0, 1, 2, 3), 
  #                      limits = c(-3.2, 3.2),
  #         labels = c("Neg bias", "", "", "NS","", "", "Bias"))+
  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
  xlab("Current b")+
  ylab("Historical b")+
  scale_fill_manual(values = c("white", "blue"))+
  coord_equal()+
  theme_classic()+
  theme(axis.text = element_text(size =10), 
        axis.title = element_text(size = 16), 
        legend.title.align = 0.05, 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14))
p

```

```{r b90, echo=FALSE, fig.cap="90% credible interval"}

# INSERT simulation and plot code
sims$a.f <- as.factor(sims$a)
sims$sig <- ifelse(sims$L90 <0 & sims$U90 >0 , "ns", 1) 
sims$sig <- ifelse(sims$L90 >0 & sims$U90 >0 , 
                                    "pos", sims$sig) 
sims$sig <- ifelse(sims$L90 <0 & sims$U90 <0 , 
                                    "neg", sims$sig) 
sims$sig.n <- 0
sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)

p = ggplot()+
  geom_tile(data = sims, aes(x = d, y = d_h,
                        fill = as.factor(sig.n)),
            color="black")+
  facet_grid(h~a)+
  # scale_fill_gradient2(expression(hat(d)), 
  #                      low = "darkred", mid = "white", 
  #                      high = "blue", 
  #                      breaks = c(-3, -2, -1, 0, 1, 2, 3), 
  #                      limits = c(-3.2, 3.2),
  #         labels = c("Neg bias", "", "", "NS","", "", "Bias"))+
  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
  xlab("Current b")+
  ylab("Historical b")+
  scale_fill_manual(values = c("white", "blue"))+
  coord_equal()+
  theme_classic()+
  theme(axis.text = element_text(size =10), 
        axis.title = element_text(size = 16), 
        legend.title.align = 0.05, 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14))
p
```

```{r b95,  echo=FALSE, fig.cap="95% credible interval"}

# INSERT simulation and plot code
sims$a.f <- as.factor(sims$a)
sims$sig <- ifelse(sims$L95 <0 & sims$U95 >0 , "ns", 1) 
sims$sig <- ifelse(sims$L95 >0 & sims$U95 >0 , 
                                    "pos", sims$sig) 
sims$sig <- ifelse(sims$L95 <0 & sims$U95 <0 , 
                                    "neg", sims$sig) 
sims$sig.n <- 0
sims$sig.n <- ifelse(sims$sig=="pos", 1, sims$sig.n)
sims$sig.n <- ifelse(sims$sig=="neg", 1, sims$sig.n)

p = ggplot()+
  geom_tile(data = sims, aes(x = d, y = d_h,
                        fill = as.factor(sig.n)),
            color="black")+
  facet_grid(h~a)+
  # scale_fill_gradient2(expression(hat(d)), 
  #                      low = "darkred", mid = "white", 
  #                      high = "blue", 
  #                      breaks = c(-3, -2, -1, 0, 1, 2, 3), 
  #                      limits = c(-3.2, 3.2),
  #         labels = c("Neg bias", "", "", "NS","", "", "Bias"))+
  scale_x_continuous(breaks = c(-3:3), expand = c(0,0))+
  scale_y_continuous(breaks = c(-3:3), expand = c(0,0))+
  xlab("Current b")+
  ylab("Historical b")+
  scale_fill_manual(values = c("white", "blue"))+
  coord_equal()+
  theme_classic()+
  theme(axis.text = element_text(size =10), 
        axis.title = element_text(size = 16), 
        legend.title.align = 0.05, 
        legend.title = element_text(size = 16), 
        legend.text = element_text(size = 14))
p
```

In these figures, _historical b_ denotes the bias parameter for the generated historical data, and current $b$ means the bias parameter for the current data. The values on the right side represent the size of the historical data. The values on the top of the plots represent the power prior weight parameter $a$. Blue means the credible interval for $b$ excludes zero.

From the plot, our model can detect strong bias and bias, when the bias parameter being -3,-2,2,3. When bias is moderate ($b = \{1, -1\}$), the model can detect it if the historical data has a bias parameter in the same direction. If the bias parameter of historical data is in an opposite direction, the model may provide a wrong estimate of current bias parameter.

# The App Prototype

The posterior probability distribution is estimated with `r R.version$version.string'[@R]. The app was built using the shiny package [@shiny].

# Discussion

# Conclusion

# CRediT Author Statement

Sachin S. Pandya: Conceptualization; Xiaomeng Li: Conceptualization; [insert]; Timothy Moore: Conceptualization; [insert]. Based on categories in https://casrai.org/credit/].

# References